---
title: "MA304_2311832"
author: "Amala Jelfa"
date: "2024-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The provided datasets are temperature and crime. The first and the foremost step is to load and understand the data. 

```{r}
# Reading the data from Excel file
df.crime <- read.csv("/Users/amala/OneDrive/Desktop/MA304_data/crime23.csv")
df.temp <- read.csv("/Users/amala/OneDrive/Desktop/MA304_data/temp2023.csv")
```

```{r}
# Creating a data frame of crime and temperature 
df.crime <- data.frame(df.crime)
df.temp <- data.frame(df.temp)
```

```{r}
# Dimension of the crime dataset
dim(df.crime)
```
There are 12 columns in the crime dataset such as category, persistent ID, date, latitude, longitude, street ID and name, location type and outcome status. 
```{r}
# Checking and analyzing the total number of missing values in crime dataset
missing_crime_val <- sapply(df.crime, function(x) sum(is.na(x)))
crime_miss <- data.frame(Col = names(missing_crime_val), Count_of_NA = missing_crime_val)
crime_miss
```

It has been found that the context column and the outcome status has missing values, which are 6878 and 677 respectively.


```{r}
# Checking and analyzing the total number of missing values in temp dataset 
missing_temp_val <- sapply(df.temp, function(x) sum(is.na(x)))
temp_miss <- data.frame(Col = names(missing_temp_val), Count_of_NA = missing_temp_val)
temp_miss
```

```{r}
# Imputing the most occurring character in the col outcome status
library(tidyverse)
getmode <- function(y) {
  uniqval <- unique(na.omit(y)) 
  uniqval[which.max(tabulate(match(y, uniqval)))]
}

df.crime$outcome_status <- replace_na(df.crime$outcome_status, getmode(df.crime$outcome_status))


View(df.crime)
df.crime <- subset(df.crime, select = -context)
```

```{r}
library(mice)
methods(mice)
df.temp <- subset(df.temp, select = -c(station_ID, PreselevHp, SnowDepcm))
impute_data <- mice(df.temp, m=10, method="pmm", seed=5000)
full_df_temp <- complete(impute_data)
summary(full_df_temp)
missing_temp_val <- sapply(full_df_temp, function(x) sum(is.na(x)))

missing_temp_val

```

```{r}
#checking for any duplicate values in the dataset 
dup_crime <- df.crime[duplicated(df.crime),]
dup_crime
dup_temp <- full_df_temp[duplicated(full_df_temp)]
dup_temp
```

There are no duplicates found in the crime and temp datasets.

```{r}
# Creating a table

crime_based <- table(df.crime$category, df.crime$outcome_status)
library(knitr)
kable(crime_based, caption = "Crimes detected based on the Category and Outcome status")

```

```{r}
#Creating a new column Month to categorize from Date 
full_df_temp$Month <- month(full_df_temp$Date, label = TRUE, abbr = TRUE)
full_df_temp
```

```{r}
install.packages(ggplotly)
# Creating a histogram using temp dataset
library(ggplot2)
library(plotly)
library(grDevices)

z <- ggplot(full_df_temp, aes(x=TemperatureCAvg, fill=Month)) + geom_histogram(stat ="bin", bins=30) + theme_classic()+ facet_wrap(~Month, ncol = 5) + theme(axis.text.x= element_text(angle = 45, vjust = 0.5, hjust = 1))
ggplotly(z, height = 400, width = 600)
 
```

```{r}
#install.packages("plotly")
library(plotly)
library(lubridate)
library(dplyr)

crime_count_pm <- df.crime %>% group_by(date) %>% summarise(crime.count =n()) %>% mutate(percentage = crime.count/ sum(crime.count)*100)
crime_count_pm 

plot_ly(crime_count_pm, labels =~date, values =~percentage, type="pie", textinfo="label+percent", insidetextorientation="radial") %>% layout(title="Number of crimes based on every month")

```

```{r}
# Result of the crime investigated according to street name
library(ggplot2)

df.crime %>%
  mutate(category=as.factor(category), long=as.factor(long), lat=as.factor(lat))

ggplot(data=df.crime) +
  geom_point(mapping = aes(x=long, y=outcome_status, color=lat))+
  geom_line(mapping = aes(x=long, y=outcome_status, color=lat))
```

```{r}
library(plotly)
full_df_temp %>%
  mutate(TemperatureCMax=as.factor(TemperatureCMax), WindkmhDir=as.factor(WindkmhDir), Month=as.factor(Month))

ggplot(data=full_df_temp)  +
  geom_bar(mapping = aes(x=TemperatureCMax, fill=WindkmhDir))

full_df_temp %>%
  group_by(TemperatureCMax) %>%
  ggplot() +
  geom_col(mapping = aes(x=TemperatureCMax, y=WindkmhDir, Month="purple"))
```


```{r}
library(ggcorrplot)
cor <- round(cor(df.temp$TemperatureCMin, df.temp$HrAvg),1)
cor_pmat(df.temp) 
GCP <- ggcorrplot(cor, hc.order = FALSE)
ggplotly(GCP,height = 400, width = 600)
df.temp$PresslevHp
```

```{r}
library(ggcorrplot)

cor_col <- full_df_temp[, c("TotClOct", "lowClOct", "WindkmhInt", "PresslevHp")]
cor_mat <- round(cor(cor_col, use = "complete.obs"), 1)
pval_mat <- cor_pmat(cor_col)

GCP <- ggcorrplot(cor_mat, hc.order = TRUE, p.mat =pval_mat, insig = "blank")
ggplotly(GCP, height = 400,width=700)
```
```{r}
library(lubridate)
library(xts)
library(forecast)
library(ggfortify)

df.dates <- ymd(full_df_temp$Date)
rain <- full_df_temp$VisKm
heavyr <- xts(data.frame(rain=rain), order.by = df.dates)
autoplot(heavyr)+theme_bw()
```
```{r}
library(xts)
prdt_r <- forecast::ma(rain, 4, centre = TRUE) 
org <- xts(data.frame(rain=rain, prdt_r), df.dates)
org <- na.omit(org)
autoplot.zoo(org,facet=NULL)+
geom_line(size=1.1) +
scale_color_manual( values = c("darkgrey","blue"), breaks = c("rain", "prdt_r"),labels = c("Rain", "Visibility in km "))+theme_bw()
```

```{r}
library(ggplot2)
library(Rcpp)
library(sf)
library(tidyverse)
library(ggmap)
library(leaflet)

crime.map <- df.crime %>%
names(crime.map) <- c("lat", "long", "n", names(crime.map)[4:5])
leaf <- crime.map %>% leaflet() %>% addTiles() %>% setView() %>% addCircleMarkers(popup=~n)
```


